Avoidance in permutations

### 2014-11-07 project2/paper/paper.tex
Addressed TODOs, removed redundant chain explanation

Since $T_n$ is only defined for even $n$, henceforth we shall refer to $T_n$ as $T_{2m}$ where $m$ is any integer greater than or equal to $0$.\todo{0 or 1 here?}

@@

Then it suffices to check that $y$ is to the right of $y-1$ and that $z$ is still to the right of $z-1$. This is easy: since $y-1$ and $z-1$ are odd numbers less than $x$ another odd number, they must be to the left of $x$, and hence to the left of $y$ and $z$.

@@

Before we do so, however, we establish a process to transform an element $a \in A$ to an element $b \in B$.

Consider a permutation $\pi$ in $T_n$ that avoids $312$. For each permutation $\pi \in T_n$ that avoids $312$, we want to show that there exists a sequence of transformation moves that transforms $\pi$ to a unique permutation $\pi'$ that avoids $321$.

If the original permutation $\pi$ avoids $321$ already, then clearly $\pi' = \pi$ avoids $321$.

Now if $\pi$ does not avoid $321$, there must exist some sub-sequence $x, y, z$ in $\pi$ in the relative ordering $321$. From Lemma \ref{swapping_lemma}, we see that swapping $y$ and $z$ in the permutation $\pi$ yields another valid permutation $\pi_1$ in $T_n$ -- note that the permutation produced after this swap has $x, y, z$ in the relative ordering $312$ instead of $321$. Note that $\pi_1$ may still have sub-sequences in the relative ordering $321$, so we may have to perform the above swapping operation again to get some permutation $\pi_2 \in T_n$ such that a sub-sequence $x_1, y_1, z_1$ in $\pi_1$ that has a relative ordering $321$ is in the relative ordering $312$ in $\pi_2$. Continuing this process produces the following chain of permutations,
$$\pi \rightarrow \pi_1 \rightarrow \pi_2 \rightarrow \ldots \rightarrow \pi_k$$
Where the last permutation in the chain $\pi_k$ does not have a single three-element sub-sequence in the relative ordering $321$, that is $\pi' = \pi_k$ is a permutation in $T_n$ that avoids $321$.\todo{why does this terminate? also, why is this necessary?}

This process is completely invertible, since we can take a permutation $\pi'$ in $T_n$ that avoids $321$, and traverse through the above chain in reverse order to obtain a permutation that avoids $312$.

@@

Let $a$ be an arbitrary permutation in $A$. Clearly, by the definition of $A$, $a$ must avoid $312$. Now, we claim that for every element $i$ in the permutation $a$, all even numbers in $a$ to the right of $i$ and less than $i$ must be in decreasing order; if this were not the case, then we would have a subsequence in the relative ordering $312$, which would be a contradiction. It is easy to see that we can make a corresponding claim about permutation $b$, that is for every element $i$ in the permutation $a$, all even numbers in $a$ to the right of $i$ and less than $i$ must be in increasing order.

\todo{this is a really long sentence}The above two claims show that there is actually an easy way to transform an arbitrary permutation $a \in A$ to a $b \in B$, and similarly, to transform an arbitrary permutation $b \in B$ to a $a \in A$ -- to obtain a permutation in $B$ from a permutation $a \in A$, for every element $i$ in the permutation $a$, all even elements in $a$ to the right of $i$ less than $i$ should be reversed (that is reordered to in increasing order); similarly to obtain a permutation in $A$ from a permutation $b \in B$, for every element $i$ in the permutation $b$, all even elements in $b$ to the right of $i$, less than $i$ should be reversed (that is reordered to in decreasing order). \todo{should have some justification here that reversing reordering these elements doesn't 'step on each other'... that is, you can always do it for some $i_1$ and then some $i_2$ afterward}
Since the mapping from $A$ to $B$ is clearly reversible, and hence invertible, we conclude that the mapping we just established between $A$ and $B$ is bijective.

>>>>

Since $T_n$ is only defined for even $n$, henceforth we shall refer to $T_n$ as $T_{2m}$ where $m$ is any integer greater than or equal to $1$.

@@

Then it suffices to check that $y$ is to the right of $y-1$ and that $z$ is still to the right of $z-1$. This is easy: since $y-1$ and $z-1$ are odd numbers less than $x$ (another odd number, by assumption), they must be to the left of $x$, and hence to the left of $y$ and $z$.

@@

@@

Let $a$ be an arbitrary permutation in $A$. Clearly, by the definition of $A$, $a$ must avoid $312$. Now, we claim that for every element $i$ in the permutation $a$, all even numbers in $a$ to the right of $i$ and less than $i$ must be in decreasing order; if this were not the case, then we would have a subsequence in the relative ordering $312$ (the element $i$ followed  by the two increasing even numbers less than $i$), which would be a contradiction. It is easy to see that we can make a corresponding claim about permutation $b$, that is for every element $i$ in the permutation $a$, all even numbers in $a$ to the right of $i$ and less than $i$ must be in increasing order.

The above two claims show that there is actually an easy way to transform an arbitrary permutation $a \in A$ to a $b \in B$, and similarly, to transform an arbitrary permutation $b \in B$ to a $a \in A$. To obtain a permutation in $B$ from a permutation $a \in A$, for every element $i$ in the permutation $a$, all even elements in $a$ to the right of $i$ less than $i$ should be reversed (that is reordered to in increasing order). Similarly to obtain a permutation in $A$ from a permutation $b \in B$, for every element $i$ in the permutation $b$, all even elements in $b$ to the right of $i$, less than $i$ should be reversed (that is reordered to in decreasing order). Note that these transformations can be expressed as multiple swaps (as described in Lemma \ref{swapping_lemma}). \todo{should have some justification here that reversing reordering these elements doesn't 'step on each other'... that is, you can always do it for some $i_1$ and then some $i_2$ afterward}
Since the mapping from $A$ to $B$ is clearly reversible, and hence invertible, we conclude that the mapping we just established between $A$ and $B$ is bijective.
